HMM Model

### 2020-09-13 manuscript/hmm.Rmd
Work towards counterfactual HMM predictions

```{r}
series_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup()

@@

hcq_m1_initial_states <- series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

@@

hcq_m1_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group
hcq_m1 <- brmshmmdata(hcq_m1_formula, series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```
```{r}
hcq_m1_epred_rect <- posterior_epred_rect(hcq_m1_fit)
hcq_m1_predicted_rect <- posterior_epred_to_predicted(hcq_m1_fit, hcq_m1_epred_rect)
hcq_m1_predicted <- posterior_rect_to_long(hcq_m1_fit, hcq_m1_predicted_rect)
```

@@

```{r}
hcq_m2_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group
hcq_m2 <- brmshmmdata(hcq_m2_formula, series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```{r}
hcq_m3_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id)
hcq_m3 <- brmshmmdata(hcq_m3_formula, series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```{r}
hcq_m4_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + took_az * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id)
hcq_m4 <- brmshmmdata(hcq_m4_formula, series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```{r}
hcq_m5_formula <-  ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id) + mo(comorbidities_sum) * rate_group
hcq_m5 <- brmshmmdata(hcq_m5_formula, series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```{r}
hcq_m6_series_data <- series_data %>% 
  group_by(.serie) %>%
  filter(any(!is.na(d_dimer))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "d_dimer", new_column_name = "peak_d_dimer", initial_value = 0) %>%
  mutate(log_peak_d_dimer = log(peak_d_dimer + 1) - 4)

hcq_m6_initial_states <- hcq_m6_series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

@@

hcq_m6 <- brmshmmdata(hcq_m6_formula, hcq_m6_series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m6_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```{r}
hcq_m7_series_data <- series_data %>% 
  group_by(.serie) %>%
  filter(any(!is.na(IL_6))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "IL_6", new_column_name = "peak_IL_6", initial_value = 1) %>%
  mutate(log_peak_IL_6 = log(peak_IL_6))

hcq_m7_initial_states <- hcq_m7_series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

@@

hcq_m7 <- brmshmmdata(hcq_m7_formula, hcq_m7_series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m7_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

>>>>

```{r}
serie_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup()

@@

hcq_m1_initial_states <- serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

@@

hcq_m1_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group
hcq_m1 <- brmshmmdata(hcq_m1_formula, serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```
```{r}

counterfactual_max_days <- 90
counterfactual_base <-  crossing(day = 0:counterfactual_max_days, patient_id = unique(serie_data$patient_id)) %>%
  inner_join(data$patient_data, by = c("patient_id")) %>% 
  left_join(wide, by = c("hospital_id", "patient_id", "day")) %>%
  mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  mutate(best_supportive_care = !is.na(best_supportive_care_from) & day >= best_supportive_care_from) %>%
  ungroup()

counterfactual_took_hcq <- counterfactual_base %>% mutate(took_hcq = TRUE)
counterfactual_no_hcq <- counterfactual_base %>% mutate(took_hcq = FALSE)

hcq_m1_counterfactual_took_hcq <- hcq_m1
hcq_m1_counterfactual_took_hcq$serie_data = counterfactual_took_hcq

hcq_m1_counterfactual_took_hcq_epred_rect <- posterior_epred_rect(hcq_m1_fit, newdata = hcq_m1_counterfactual_took_hcq)

hcq_m1_counterfactual_no_hcq <- hcq_m1
hcq_m1_counterfactual_no_hcq$serie_data = counterfactual_no_hcq

hcq_m1_counterfactual_no_hcq_epred_rect <- posterior_epred_rect(hcq_m1_fit, newdata = hcq_m1_counterfactual_no_hcq)

```{r}
hcq_m1_epred_rect <- posterior_epred_rect(hcq_m1_fit)
hcq_m1_predicted_rect <- posterior_epred_to_predicted(hcq_m1_fit, hcq_m1_epred_rect)
hcq_m1_predicted <- posterior_rect_to_long(hcq_m1_fit, hcq_m1_predicted_rect)
```

@@

```{r}
hcq_m2_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group
hcq_m2 <- brmshmmdata(hcq_m2_formula, serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```{r}
hcq_m3_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id)
hcq_m3 <- brmshmmdata(hcq_m3_formula, serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```{r}
hcq_m4_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + took_az * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id)
hcq_m4 <- brmshmmdata(hcq_m4_formula, serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```{r}
hcq_m5_formula <-  ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id) + comorbidities_sum * rate_group #mo(comorbidities_sum) * rate_group
hcq_m5 <- brmshmmdata(hcq_m5_formula, serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```{r}
hcq_m6_serie_data <- serie_data %>% 
  group_by(.serie) %>%
  filter(any(!is.na(d_dimer))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "d_dimer", new_column_name = "peak_d_dimer", initial_value = 0) %>%
  mutate(log_peak_d_dimer = log(peak_d_dimer + 1) - 4)

hcq_m6_initial_states <- hcq_m6_serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

@@

hcq_m6 <- brmshmmdata(hcq_m6_formula, hcq_m6_serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m6_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

```{r}
hcq_m7_serie_data <- serie_data %>% 
  group_by(.serie) %>%
  filter(any(!is.na(IL_6))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "IL_6", new_column_name = "peak_IL_6", initial_value = 1) %>%
  mutate(log_peak_IL_6 = log(peak_IL_6))

hcq_m7_initial_states <- hcq_m7_serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

@@

hcq_m7 <- brmshmmdata(hcq_m7_formula, hcq_m7_serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m7_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)
